---
title: "Hepatitis"
output: html_notebook
---

Variable names and data were in a separate file. Start by assigning names to variables.

```{r}
hepatitis_raw <- read.csv("C:/Users/kelse/Documents/Year_2/Rotations/Colubri/Hepatitis_data/hepatitis.csv", header = FALSE, sep = ",")

hepatitis <- hepatitis_raw

names <- c("Class", "Age", "Sex", "Steroid",  "Antivirals", "Fatigue", "Malaise", "Anorexia", "Liver_big", "Liver_firm",  "Spleen_palpable",  "Spiders", "Ascites",  "Varices", "Bilirubin", "Alk_phosphate", "Sgot",  "Albumin", "Protime",  "Histology")

names(hepatitis) <- names

head(hepatitis)
```


It looks like missing values are recorded as "?" in this dataset. Start by changing these to NA.

```{r}
library(naniar)
hepatitis <- hepatitis %>% 
  replace_with_na_all(condition = ~.x == "?")

sapply(hepatitis, function(x) sum(is.na(x))) # Display total NAs in each variable
```

Replace values of 1 and 2 with die and live, respectively (got this info from hepatitis NAMES file).

```{r}
hepatitis <- within(hepatitis, {
Class[Class == 1] <- "die"
Class[Class == 2] <- "live"
})

```

Convert character variables to factors
```{r}

hepatitis$Class <- factor(hepatitis$Class)

#Convert character variables to factors
hepatitis$Age <- as.numeric(as.character(hepatitis$Age))
hepatitis$Sex <- as.numeric(as.character(hepatitis$Sex))
hepatitis$Steroid <- as.numeric(as.character(hepatitis$Steroid))
hepatitis$Antivirals <- as.numeric(as.character(hepatitis$Antivirals))
hepatitis$Fatigue <- as.numeric(as.character(hepatitis$Fatigue))
hepatitis$Malaise <- as.numeric(as.character(hepatitis$Malaise))
hepatitis$Anorexia <- as.numeric(as.character(hepatitis$Anorexia))
hepatitis$Liver_big <- as.numeric(as.character(hepatitis$Liver_big))
hepatitis$Liver_firm <- as.numeric(as.character(hepatitis$Liver_firm))
hepatitis$Spleen_palpable <- as.numeric(as.character(hepatitis$Spleen_palpable))
hepatitis$Spiders <- as.numeric(as.character(hepatitis$Spiders))
hepatitis$Ascites <- as.numeric(as.character(hepatitis$Ascites))
hepatitis$Varices <- as.numeric(as.character(hepatitis$Varices))
hepatitis$Bilirubin <- as.numeric(as.character(hepatitis$Bilirubin))
hepatitis$Alk_phosphate <- as.numeric(as.character(hepatitis$Alk_phosphate))
hepatitis$Sgot <- as.numeric(as.character(hepatitis$Sgot))
hepatitis$Albumin <- as.numeric(as.character(hepatitis$Albumin))
hepatitis$Protime <- as.numeric(as.character(hepatitis$Protime))
hepatitis$Histology <- as.numeric(as.character(hepatitis$Histology))

```

Subset binary variables into new data frame "hep_binary"

```{r}
library(tidyverse)
hep_binary <- hepatitis %>%
  select(6:14,20)
```

Write a function to determine overall incidence of each binary variable. Sum values equal to 1 for each column, then divide by total number of cases (155).

```{r}
Overall_incidence_func <- function(x) {
  (length(which(x==1))) / 155 * 100
}

Overall_incidence <- sapply(hep_binary, Overall_incidence_func)

```

TRY SOMETHING ELSE

```{r}
a <- sapply(hep_binary, function(x) length(which(x==1)))
b <- round(Overall_incidence, 0)

for (i in hep_binary) {
  OI <- paste(a,"/155"," (",b,"%)", sep="")
}
```

Calculate survival. Overall 123/155 survived. 

total # cases where symptom occurred and outcome = survival / 123

```{r}

#write a function to create a table of symptoms summarized by live or die and then extract the second element of the table corresponding to cases where patient experienced symptom and survived
select_survival <- function(col) {
  (with(hepatitis, table(Class, col))[2]) / 123 * 100
}

#apply function to all binary variables and save in new variable "Survival"
Survival <- sapply(hep_binary, select_survival)
```


Calculate death. Overall 32/155 died. 

total # cases where symptom occurred and outcome = death / 32

```{r}

#write a function to create a table of symptoms summarized by live or die and then extract the second element of the table corresponding to cases where patient experienced symptom and survived
select_death <- function(col) {
  (with(hepatitis, table(Class, col))[3]) / 32 * 100
}

#apply function to all binary variables and save in new variable "Survival"
Death <- sapply(hep_binary, select_death)
```

Create a summary table of outcomes for the data 
```{r}
table_func <- function(col) {
  (with(hepatitis, table(Class, col)))
}

group <- c("die, symptom(+)", "live, symptom(+)", "die, symptom(-)", "live, symptom(-)")

summary_table <- sapply(hep_binary, table_func)
dimnames(summary_table) <- list("Group" = group, "Symptom" = names(hep_binary))

```

Create survival outcome table for each binary variable and save in a list.
```{r}
#write function to create individual outcome tables for each binary variable and save table as list named test4
table_func2 <- function(col) {
  (with(hepatitis, table(col, Class)))
}

test4 <- map(hep_binary, table_func2)
```

Format survival data as a string 

```{r}
a <- map(test4, 3)
b <- round(Survival,0)

for (i in hep_binary) {
  Survival <- paste(a,"/123"," (",b,"%)", sep="")
}

```

Format death data as a string 

```{r}
a <- map(test4, 1)
b <- round(Death,0)


for (i in hep_binary) {
  Death <- paste(a,"/32"," (",b,"%)", sep="")
}
Death 
```

Calculate odds ratio for each binary variable. 

```{r}
library(questionr)

#add odds ratio for each variable to the summary table df
test6 <- map(test4, odds.ratio)
OR <- map(test6, "OR") 
OR <- round(data.frame(unlist(OR)), 1)

#add p-value for each odds ratio to the summary table df
p_value <- map(test6, "p") 
p <- round(data.frame(unlist(p_value)), 2)

```

```{r}
names <- list(names(hep_binary))
df <- data.frame(OI, Survival, Death, p, OR)
names(df) <- c("Overall Incidence", "Survival", "Death", "p value", "Odds ratio")
df
```

