---
title: "Hepatitis"
output: html_notebook
---
Load packages
```{r}
library(naniar)
library(tidyverse)
library(stringr)
library(DescTools)
```

Variable names and data were in a separate file. Start by assigning names to variables.

```{r}
rm(list=ls())
hepatitis_raw <- read.csv("C:/Users/kelse/Documents/Year_2/Rotations/Colubri/Hepatitis_data/hepatitis.csv", header = FALSE, sep = ",")
hepatitis <- hepatitis_raw
names <- c("Class", "Age", "Sex", "Steroid",  "Antivirals", "Fatigue", "Malaise", "Anorexia", "Liver_big", "Liver_firm",  "Spleen_palpable",  "Spiders", "Ascites",  "Varices", "Bilirubin", "Alk_phosphate", "Sgot",  "Albumin", "Protime",  "Histology")
names(hepatitis) <- names
```


It looks like missing values are recorded as "?" in this dataset. Start by changing these to NA.

```{r}
hepatitis <- hepatitis %>% 
  replace_with_na_all(condition = ~.x == "?")
```

Replace values of 1 and 2 with die and live, respectively (got this info from hepatitis NAMES file).

```{r}
hepatitis <- within(hepatitis, {
Class[Class == 1] <- "die"
Class[Class == 2] <- "live"
})
```

Convert character variables to factors
```{r}
hepatitis$Class <- factor(hepatitis$Class)

#Convert character variables to factors
hepatitis$Age <- as.numeric(as.character(hepatitis$Age))
hepatitis$Sex <- as.numeric(as.character(hepatitis$Sex))
hepatitis$Steroid <- as.numeric(as.character(hepatitis$Steroid))
hepatitis$Antivirals <- as.numeric(as.character(hepatitis$Antivirals))
hepatitis$Fatigue <- as.numeric(as.character(hepatitis$Fatigue))
hepatitis$Malaise <- as.numeric(as.character(hepatitis$Malaise))
hepatitis$Anorexia <- as.numeric(as.character(hepatitis$Anorexia))
hepatitis$Liver_big <- as.numeric(as.character(hepatitis$Liver_big))
hepatitis$Liver_firm <- as.numeric(as.character(hepatitis$Liver_firm))
hepatitis$Spleen_palpable <- as.numeric(as.character(hepatitis$Spleen_palpable))
hepatitis$Spiders <- as.numeric(as.character(hepatitis$Spiders))
hepatitis$Ascites <- as.numeric(as.character(hepatitis$Ascites))
hepatitis$Varices <- as.numeric(as.character(hepatitis$Varices))
hepatitis$Bilirubin <- as.numeric(as.character(hepatitis$Bilirubin))
hepatitis$Alk_phosphate <- as.numeric(as.character(hepatitis$Alk_phosphate))
hepatitis$Sgot <- as.numeric(as.character(hepatitis$Sgot))
hepatitis$Albumin <- as.numeric(as.character(hepatitis$Albumin))
hepatitis$Protime <- as.numeric(as.character(hepatitis$Protime))
hepatitis$Histology <- as.numeric(as.character(hepatitis$Histology))

```

Subset binary variables into new data frame "hep_binary"

```{r}
hep_binary <- hepatitis %>%
  select(6:14,20)

hep_binary_class <- hepatitis %>%
  select(1,6:14,20)

```

Write a function to determine overall incidence of each binary variable. Sum values equal to 2 for each column, then divide by total number of cases excluding NA

```{r}
total_na <- sapply(hep_binary, function (x) sum(is.na(x)))
total <- 155 - total_na

#write function to calculate total # of 2 corresponding to yes
Overall_incidence_func <- function(x) {
  (length(which(x==2)))
} 

overall <- sapply(hep_binary, Overall_incidence_func)

Overall_incidence <- overall / total * 100
Overall_incidence
```

Save overall incidence data in a string variable to report later

```{r}
a <- sapply(hep_binary, function(x) length(which(x==2)))
b <- round(Overall_incidence, 0)

for (i in hep_binary) {
  OI <- paste(a,"/",total," (",b,"%)", sep="")
}

OI
```

Calculate total # of cases that with outcome survival/death for each variable, excluding NA
```{r}
#survival
survival <- hepatitis %>%
  filter(Class == "live") %>%
  select(6:14,20)

missing_surv <- sapply(survival, function (x) sum(is.na(x))) # calculate total NA for each variable
n_survival <- 123 - missing_surv

#death 
death <- hepatitis %>%
  filter(Class == "die") %>%
  select(6:14,20)

missing_death <- sapply(death, function (x) sum(is.na(x))) # calculate total NA for each variable
n_death <- 32 - missing_death
```

Calculate survival. 

```{r}
#write a function to create a contingency table for each binary variable
table_func <- function(col) {
  (with(hepatitis, table(col, Class)))
}

cont_table <- map(hep_binary, table_func)
cont_table <- map(cont_table, function (x) Rev(x, 1)) #reverse contingency table so values where symptom = yes(2) are on the top row
cont_table
```

Extract the third (top right) element of the table corresponding to cases where patient experienced symptom and survived
Extract the first (top left) element of the table corresponding to cases where patient experienced symptom and died

```{r}
exposed_live <- map(cont_table, 3)
exposed_die <- map(cont_table, 1)

survival <- as.numeric(exposed_live) / n_survival * 100
death <- as.numeric(exposed_die) / n_death * 100
survival_percent <- round(survival,0)
death_percent <- round(death,0)
```

Save survival and death data in a string to be reported later.

```{r}
for (i in hep_binary) {
  surv_string <- paste(exposed_live,"/",n_survival," (",survival_percent,"%)", sep="")
}

for (i in hep_binary) {
  death_string <- paste(exposed_die,"/",n_death," (",death_percent,"%)", sep="")
}
```

Calculate odds ratio for each binary variable. 

```{r}
#add odds ratio for each variable to the summary table df
odds_ratio <- map(cont_table, odds.ratio)
OR <- map(odds_ratio, "OR") 
OR <- round(data.frame(unlist(OR)), 2)
odds_ratio

#add p-value for each odds ratio to the summary table df
p_value <- map(odds_ratio, "p") 
p <- round(data.frame(unlist(p_value)), 2)
```

Create summary data frame for all binary variables.

```{r}
names <- list(names(hep_binary))
df <- data.frame(OI, surv_string, death_string, p, OR)
names(df) <- c("Overall Incidence", "Survival", "Death", "p value", "Odds ratio")
df
```

```{r}
fat <- hepatitis[1,6]
head(fat)
```

