---
title: "Hepatitis"
output: html_notebook
---

Variable names and data were in a separate file. Start by assigning names to variables.

```{r}
hepatitis_raw <- read.csv("C:/Users/kelse/Documents/Year_2/Rotations/Colubri/Hepatitis_data/hepatitis.csv", header = FALSE, sep = ",")

hepatitis <- hepatitis_raw

names <- c("Class", "Age", "Sex", "Steroid",  "Antivirals", "Fatigue", "Malaise", "Anorexia", "Liver_big", "Liver_firm",  "Spleen_palpable",  "Spiders", "Ascites",  "Varices", "Bilirubin", "Alk_phosphate", "Sgot",  "Albumin", "Protime",  "Histology")

names(hepatitis) <- names

head(hepatitis)
```


It looks like missing values are recorded as "?" in this dataset. Start by changing these to NA.

```{r}
library(naniar)
hepatitis <- hepatitis %>% 
  replace_with_na_all(condition = ~.x == "?")

sapply(hepatitis, function(x) sum(is.na(x))) # Display total NAs in each variable
```

Replace values of 1 and 2 with die and live, respectively (got this info from hepatitis NAMES file).

```{r}
hepatitis <- within(hepatitis, {
Class[Class == 1] <- "die"
Class[Class == 2] <- "live"
})
head(hepatitis, 10)

```


### Use transcan to impute missing values
```{r}
htrans <- #creates a transcan object with results of transcan imputation
  transcan(~ Steroid + Fatigue + Malaise + Anorexia + Liver_big + Liver_firm + 
             Spleen_palpable + Spiders + Ascites + Varices + Bilirubin + 
             Alk_phosphate + Sgot + Albumin + Protime,
           imputed=TRUE, transformed=TRUE, 
           data=hepatitis, pl=FALSE, pr=FALSE)

#impute creates a new data set with all values of imputed variables (designated with *)
hep_imp <- impute(htrans, data=hepatitis, list.out=TRUE)

```

Next add imputed values back to original data set.
```{r}

#all.vars returns a character vector containing all names in the expression. It seems this would be the
#same as making a vector like c("Steroid", "Fatigue", ...). all.vars() is a little more streamlined
#because you don't have to type quotes.
NAvars <- all.vars(~ Steroid + Fatigue + Malaise + Anorexia + Liver_big + Liver_firm + Spleen_palpable + Spiders + Ascites + Varices + Bilirubin + Alk_phosphate + Sgot + Albumin + Protime) 

for(x in NAvars) hepatitis[[x]] <- hep_imp[[x]] #assign imputed values in imp back to original dataset

sapply(hepatitis, function(x) sum(is.na(x))) #check that all NAs have been replaced

```

### Build a logistic regression model 
Convert outcome variable to factor
Convert other variables to numeric

When I try the following code I get an error that says "NAs introduced by coercion" and all my data becomes NA

```{r}
hepatitis_1 <- hepatitis

for(i in 2:20) {
 hepatitis_1[, i] <- as.numeric(as.character(hepatitis[, i]))
}
```

Next I tried using sapply to do the same thing, but this didn't seem to work either. Bilirubin, Albumin, and some other variables didn't convert correctly.

```{r}
hep <- sapply(hepatitis, as.character)
hep <- sapply(hepatitis, as.numeric)

```

```{r}
head(hepatitis)
```

```{r}

hepatitis$Class <- factor(hepatitis$Class)

#Convert character variables to factors
hepatitis$Age <- as.numeric(as.character(hepatitis$Age))
hepatitis$Sex <- as.numeric(as.character(hepatitis$Sex))
hepatitis$Steroid <- as.numeric(as.character(hepatitis$Steroid))
hepatitis$Antivirals <- as.numeric(as.character(hepatitis$Antivirals))
hepatitis$Fatigue <- as.numeric(as.character(hepatitis$Fatigue))
hepatitis$Malaise <- as.numeric(as.character(hepatitis$Malaise))
hepatitis$Anorexia <- as.numeric(as.character(hepatitis$Anorexia))
hepatitis$Liver_big <- as.numeric(as.character(hepatitis$Liver_big))
hepatitis$Liver_firm <- as.numeric(as.character(hepatitis$Liver_firm))
hepatitis$Spleen_palpable <- as.numeric(as.character(hepatitis$Spleen_palpable))
hepatitis$Spiders <- as.numeric(as.character(hepatitis$Spiders))
hepatitis$Ascites <- as.numeric(as.character(hepatitis$Ascites))
hepatitis$Varices <- as.numeric(as.character(hepatitis$Varices))
hepatitis$Bilirubin <- as.numeric(as.character(hepatitis$Bilirubin))
hepatitis$Alk_phosphate <- as.numeric(as.character(hepatitis$Alk_phosphate))
hepatitis$Sgot <- as.numeric(as.character(hepatitis$Sgot))
hepatitis$Albumin <- as.numeric(as.character(hepatitis$Albumin))
hepatitis$Protime <- as.numeric(as.character(hepatitis$Protime))
hepatitis$Histology <- as.numeric(as.character(hepatitis$Histology))

head(hepatitis)

```


Now that variables are in correct format, try glm() again.
```{r}
glm.fit <- glm(Class ~ Age + Sex + Steroid + Antivirals + Fatigue + Malaise + Anorexia + Liver_big + Liver_firm + Spleen_palpable + Spiders + Ascites + Varices + Bilirubin + Alk_phosphate + Sgot + Albumin + Protime + Histology, data = hepatitis, family = binomial)

summary(glm.fit)
```

Next make predictions on the training data that I used to fit the model. Set type = response to change to the binary scale of the response variable. Assign the results to glm.probs to save a table of probabilities. 

#'*QUESTION: Probability table contains what the model would have predicted based on this training set. I think next I should divide the data into training and test subsets.*
```{r}
glm.probs <- predict(glm.fit,type = "response")
glm.probs[1:10]
```

```{r}
glm.pred <- ifelse(glm.probs > 0.5, "live", "die")

attach(hepatitis) ## this returned an error message
table(glm.pred,Class)
```

```{r}
mean(glm.pred == Class)
```

The model was correct ~90% of the time, try to make it better by dividing into training and test sets.

Use createDataPartition() to maintain the balance of the factor classes in the training and test data sets. 
#'*I'm not sure if I need to use upsampling here or not.*


```{r}
#using this code from a tutorial
library(caret)
'%ni%' <- Negate('%in%') 
options(scipen=999) 

# Prep Training and Test data
set.seed(100)
trainDataIndex <- createDataPartition(hepatitis$Class, p=0.7, list = F)  # 70% training data, list=F puts the results in a matrix
trainData <- hepatitis[trainDataIndex, ]
testData <- hepatitis[-trainDataIndex, ]

summary(trainData$Class)
summary(testData$Class)
str(testData)
```

Next fit the model to the training data (110 cases)
```{r}
logitmod <- glm(Class ~ Age + Sex + Steroid + Antivirals + Fatigue + Malaise + Anorexia + Liver_big + Liver_firm + Spleen_palpable + Spiders + Ascites + Varices + Bilirubin + Alk_phosphate + Sgot + Albumin + Protime + Histology, data = trainData, family = binomial)

summary(logitmod)
```

Use this model to predict the response on the testData. The pred variable contains the probability that the observation is live for each case.

```{r}
pred <- predict(logitmod, newdata = testData, type = "response")
```

```{r}
y_pred_num <- ifelse(pred > 0.5, "live", "die")
y_pred <- factor(y_pred_num, levels=c("die","live"))
y_act <- testData$Class
mean(y_pred == y_act) #another way to generate confusion matrix
summary(y_act)
```
 The model has an accuracy rate of 87%.
 
 Next create confusion matrix. It shows:
 Accuracy of 87% which is TP / (TP + FP) #'*ANDRES: wouldn't this be 3/(3+0)...I am confused here*
 Sensitivity of 33% which is TP/P or 3/9
 Specificity of 100% which is TN/N or 36/36

```{r}
library(caret)
confusionMatrix(y_pred, testData[["Class"]])
```

Now plot an ROC curve - code from https://rpubs.com/dtime/672367
#'*ANDRES: I used code from the url above, what are the rules regarding citations/crediting sources you borrow code from?*

```{r}
library(ROCR)
roc_pred <- prediction(predictions = pred, labels = testData$Class)
roc_perf <- performance(roc_pred, "tpr", "fpr")
plot(roc_perf,
     colorize = TRUE)
```

Now calculate the AUC
```{r}
auc_ROCR <- performance(roc_pred, measure = "auc")

#one way to calculate the AUC
auc_method1 <- auc_ROCR@y.values[[1]]
auc_method1
```

And another way to calculate the AUC

```{r}
auc_method2 <- unlist(slot(auc_ROCR, "y.values"))
auc_method2
```




